---
title: "Statistical Distributions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Poisson Distribution
The poisson distribution is controlled by a single parameter $\lambda$. 
The poisson probability mass function is $P(X=x) = \frac{\lambda^k e^-\lambda}{k!}$, where $k=0,1,2...$.
We are interested in how the distribution changes as $\lambda$ changes.  
<br>

For a start, let's set $\lambda = 1$ and see what the PMF looks like. We see that 0 and 1 are equally likely, and the likelihood of higher values quickly declines to 0 thereafter.    

```{r cars, echo=F}
library(ggplot2); library(scales)
k <- 0:10
lambda <- 1
df <- data.frame(p = exp(-lambda) * lambda^k / factorial(k), k=k)

ggplot(df, aes(x=k, y=p)) + geom_line() + geom_point() + theme_bw() +
  scale_y_continuous(breaks=seq(0,1,0.1), labels=percent_format()) +
  scale_x_continuous(breaks=0:10) +
  labs(title="Probability Mass Function of Poisson", subtitle="Lambda = 1")

```

<br>

Now let's try varying $\lambda = 1...,10$ and see what the PMF looks like. You can see that the curve gradually shifts towards the right as the $\lambda$ or rate parameter increases. The curve for $\lambda = 5$ seems symmetrically distributed around $k = 5$, which is intuitive as the expected value of a poisson distribution is also its rate, or $EX = \lambda$. 
```{r, echo=F}
library(ggplot2); library(scales)
lambda <- seq(1, 10, 1)
k <- 0:20
df <- data.frame()
for (l in lambda) {
  p = exp(-l) * l^k / factorial(k)
  temp = data.frame(p=p, k=k, lambda=l)
  df <- rbind(df, temp)
}

ggplot(df, aes(x=k, y=p, col=factor(lambda), group=lambda)) + 
  geom_line() + geom_point() + theme_bw() +
  scale_y_continuous(breaks=seq(0,1,0.1), labels=percent_format()) +
  scale_x_continuous(breaks=seq(0,20,2)) +
  scale_color_brewer(palette=5, type="div", name="Lambda Value") +
  labs(title="Probability Mass Function of Poisson Distribution", 
       subtitle="Lambda = 1 to 10") 

```

<br>

## Normal Distribution
The standard normal distribution has probability density function of $f(z) = \frac{1}{\sqrt2\pi}e^{\frac{-z^2}{2}}$.
<br>

Let's look at what the PDF looks like.  
```{r, echo=F}
library(ggplot2); library(scales)
df <- data.frame(z=seq(-5,5,0.01))
df$pdf <- 1/sqrt(2 * pi) * exp(-(df$z)^2/2)

ggplot(df, aes(x=z, y=pdf)) + 
  geom_line(size=1.5) + theme_bw() +
  scale_y_continuous(breaks=seq(0,5,0.1), labels=percent_format()) +
  scale_x_continuous(breaks=-5:5) +
  labs(title="Probability Density Function of Standard Normal")

```

